### 2015-06-11 testing-code/check-sim.Rmd
.

# Compare simulation to expected counts under true parameters

```{r, echo=FALSE}
# Should load: simulation results; generator matrix
library(pander)
library(ggplot2)
panderOptions("table.split.table", Inf)  # Let tables become arbitrarily wide rather than wrapping.
source("../context-inference-fns.R")
source("../sim-context-fns.R")
top_n_bottom <- function(df, k) { # Take the top and bottom k rows
    n = nrow(df)
    if(2*k < n) {
        df[c(1:k,(n-(k-1)):n), ]
    }
    else {
        df
    }
}
counts_table <- function(in_longwin, in_shortwin, half_n_show) {
    if ( in_longwin > longwin(genmatrix) || in_shortwin > longwin(genmatrix) ) {
        cat("Window size of counts too small.")
        return(NULL)
    }
    else {
        observed <- projectcounts( counttab, new.leftwin=floor((in_longwin-in_shortwin)/2), new.shortwin=in_shortwin, new.longwin=in_longwin, overlapping=TRUE )
        expected <- projectcounts( expected, new.leftwin=floor((in_longwin-in_shortwin)/2), new.shortwin=in_shortwin, new.longwin=in_longwin, overlapping=TRUE )
        obsexp <- cbind( as.matrix(observed@counts), as.matrix(expected@counts) )
        colnames(obsexp) <- c( paste("obs:",colnames(observed),sep=''), paste("exp:",colnames(observed),sep='') )
        obsexp <- obsexp[, rbind(1:ncol(observed),ncol(observed)+(1:ncol(observed)))]
        pandoc.table(top_n_bottom(obsexp, half_n_show), style="rmarkdown")
    }
    return(invisible(obsexp))
}
longpats <- getpatterns(longwin(genmatrix),simseq.config$bases)
counttab <- counttrans.list( list(longpats,longpats), simseqs=simseqs[c("root","tip")], leftwin=0, bases=simseq.config$bases )
initcounts <- rowSums(counttab)
time <- as.numeric(simseq.opt$tlen)
expected <- predictcounts(longwin=longwin(genmatrix), shortwin=longwin(genmatrix), leftwin=0, initcounts=initcounts, 
        mutrates=time*simseq.config$tip$mutrates, selcoef=simseq.config$tip$selcoef, genmatrix=genmatrix, params=simseq.config$tip$fixfn.params )
```

# Simulated parameter values
```{r, echo=FALSE, results="asis"}
params <- as.data.frame( matrix( c(
        time,
        simseq.config$tip$mutrates,
        simseq.config$tip$selcoef,
        simseq.config$tip$fixfn.params
    ), nrow=1 ) )
colnames(params) <- gsub("|","\\|",c(
        "time",
        mutnames(simseq.config$tip$mutpats,pretty=TRUE),
        selnames(simseq.config$tip$selpats,pretty=TRUE),
        names(simseq.config$tip$fixfn.params)
    ), fixed=TRUE)
pandoc.table(params, style="rmarkdown")
```

```{r, echo=FALSE}
cat("window size:", longwin(genmatrix),
    "\ntotal counts:", sum(counttab@counts),
    "\ntime:", simseq.opt$tlen,
    "\nseed:", simseq.opt$seed,
    "\n"
    )
```

## residuals for 2-2 Tmers
```{r, echo=FALSE, results="asis"}
counts_table(in_longwin=2, in_shortwin=2, half_n_show=20)
```

## residuals for 3-1 Tmers
```{r, results="asis"}
counts_table(in_longwin=3, in_shortwin=1, half_n_show=20)
```

## residuals for 9-5 Tmers
```{r, results="asis"}
counts_table(in_longwin=9, in_shortwin=5, half_n_show=20)
```