### 2015-07-29 simulation.Rmd
added hessian to output

# Results from fitting procedure

*Invocation:*
```{r, echo=FALSE}
library(pander)
library(ggplot2)
panderOptions("table.split.table", Inf)  # Let tables become arbitrarily wide rather than wrapping.
source("context-inference-fns.R",chdir=TRUE)
top_n_bottom <- function(df, k) { # Take the top and bottom k rows
    n = nrow(df)
    if(2*k < n) {
        df[c(1:k,(n-(k-1)):n), ]
    }
    else {
        df
    }
}
counts_table <- function(in_longwin, in_shortwin, half_n_show) {
    if ( in_longwin > longwin(model@counts) || in_shortwin > shortwin(model@counts) ) {
        cat("Window size of counts too small.")
        return(NULL)
    }
    else {
        in_leftwin <- floor((in_longwin-in_shortwin)/2)
        resids <- computeresids(model,
                in_longwin=in_longwin,
                in_shortwin=in_shortwin,
                in_leftwin=in_leftwin,
                counts=model@counts, 
                overlapping=TRUE)
        true.expected <- projectcounts(
                fitted( true.model, longwin=longwin(model), shortwin=shortwin(model), leftwin=leftwin(model) ),
                in_leftwin, in_shortwin, in_longwin, 
                overlapping=TRUE )
        resids$sim.expected <- true.expected@counts[ cbind(resids$inpat,resids$outpat) ]
        pandoc.table(top_n_bottom(resids, half_n_show), style="rmarkdown")
        return(invisible(resids))
    }
}
finite_range <- function (...) { x <- unlist( list(...) ); range(x[is.finite(x)&!is.na(x)]) }
counts_plot <- function (resids) {
    if (NROW(resids)>0) { 
        hist(resids$z, xlab="z-score", ylab='', main="z score") 
        outfac <- factor(resids$outpat)
        with(resids, {
                plot(observed, expected, xlab='observed', ylab='expected', col=rainbow(nlevels(outfac))[as.numeric(outfac)], pch=20, ylim=finite_range(expected,sim.expected) ) 
                # points(observed, sim.expected, xlab='observed', ylab='expected', col=adjustcolor(rainbow(nlevels(outfac))[as.numeric(outfac)],.5), pch=1, cex=.75 ) 
                segments(x0=observed, y0=expected, y1=sim.expected, col=adjustcolor(rainbow(nlevels(outfac))[as.numeric(outfac)],.5) )
            } )
        abline(0,1)
        with(resids, {
                plot(observed, expected, xlab='observed', ylab='expected', log='xy', col=rainbow(nlevels(outfac))[as.numeric(outfac)], pch=20, ylim=pmax(1e-4,finite_range(expected,sim.expected)) ) 
                segments(x0=observed, y0=expected, y1=sim.expected, col=adjustcolor(rainbow(nlevels(outfac))[as.numeric(outfac)],.5) )
            } )
        abline(0,1)
    }
}
counttab <- counts(model)
time <- as.numeric(simseq.opt$tlen)
timevec <- c( rep(as.numeric(simseq.opt$tlen),nmuts(model)), rep(1,length(coef(model))-nmuts(model)) )
true.model <- model
true.model@mutrates <- simseq.config$tip$mutrates*time
true.model@selcoef <- simseq.config$tip$selcoef
if (length(true.model@params)>0) { true.model@params <- unlist(simseq.config$tip$fixfn.params) }
mr_compare <- data.frame( 
    fit=coef(model)/timevec,
    simulated=coef(true.model)/timevec,
    fixed=!model@results$use.par,
    stringsAsFactors=FALSE )
cat(model@invocation,"\n")
```

```{r summary, echo=FALSE, results="asis"}
cat(
    "\n*  long window size:", longwin(model@counts),
    "\n*  short window size:", shortwin(model@counts),
    "\n*  left window size:", leftwin(model@counts),
    "\n*  total counts:", sum(counttab),
    "\n*  time:", simseq.opt$tlen,
    "\n*  seed:", formatC(simseq.opt$seed,format="d"),
    "\n"
    )
```

## Result of fitting operation
```{r convergence, echo=FALSE, results="asis"}
if (class(model)=="context") {
    cat("\n* convergence: ", model@results$convergence,
        "\n* number of iterations: ", model@results$counts,
        "\n* message: ", model@results$message,
        "\n" )
} else if (class(model)=="contextMCMC") {
    cat(
        "\n* number of batches: ", model@results$nbatch,
        "\n* batch length: ", model@results$blen,
        "\n* acceptance rate: ", model@results$accept,
        "\n* run time: ", model@results$time['elapsed'],
        "\n" )
}
```

## Fitted parameter values
```{r fitted, echo=FALSE}
pander(mr_compare)
mr_compare$names <- factor(names(coef(model)),levels=names(coef(model)))  # factor explicitly to get ggplot to put bars in order
```

The corresponding plot, where bars are the fit values, and points are the simulated values:
```{r fit_plot, echo=FALSE, results="asis", fig.align='center'}
param_log <- log(abs(mr_compare$simulated)+1e-4)
param_sizes <- cut( param_log, breaks=seq(min(param_log), by=2, length.out=max(2,1+ceiling(diff(range(param_log))/2))), include.lowest=TRUE )
for (k in unique(param_sizes)) {
    print( ggplot(mr_compare[param_sizes==k,], aes(x=names)) + xlab("") +
        geom_bar(aes(y=fit), stat = "identity", fill="white", color="gray") +
        geom_point(aes(y=simulated)) +
        theme(axis.text.x = element_text(angle = 90),
              axis.title.y = element_blank())
      )
}
```

The profile likelihood surface about the fitted and simulated values (red are the values simulated under, black is the fitted values):
```{r likelihood_surface, echo=FALSE, results="asis", fig.align='center'}
f <- function (x) { model@likfun((x*timevec)[model@results$use.par]) }
for (k in (1:nrow(mr_compare))[model@results$use.par]) {
    pvals <- seq( .9*min(mr_compare[k,c("fit","simulated")]), 1.1*max(mr_compare[k,c("fit","simulated")]), length.out=10 )
    inds <- 1:nrow(mr_compare)
    inds[k] <- 1+nrow(mr_compare)
    lvals <- sapply( pvals, function (x) { f( c(mr_compare$fit,x)[inds] ) } )
    l2vals <- sapply( pvals, function (x) { f( c(mr_compare$simulated,x)[inds] ) } )
    plot( pvals, lvals, type='b', main=paste("likelihood surface for", mr_compare$name[k]), ylab="log-likelihood", xlab='value', ylim=finite_range(lvals,l2vals) )
    lines( pvals, l2vals, col='red', type='b' )
    abline(v=mr_compare[k,c("fit","simulated")], col=c("black","red"))
}
```

Here is the Hessian of the likelihood (obtained by numerical differentiation)
```{r likelihood_hessian, echo=FALSE}
# Hessian of the likelihood surface
likfun.hess <- hessian(function(x)fit.model@likfun(x*timevec[fit.model@results$use.par]),method.args=list(eps=min(coef(model)/10)),x=coef(model)[fit.model@results$use.par])
colnames( likfun.hess ) <- rownames( likfun.hess ) <- names(coef(fit.model))[fit.model@results$use.par]
pander(likfun.hess)
```
and its inverse (the asymptotic covariance matrix)
```{r likelihood_covariance, echo=FALSE}
# asymptotic covariance matrix
est.covmat <- ginv(likfun.hess)
pander(est.covmat)
```
and the corresponding correlation matrix
```{r likelihood_covariance, echo=FALSE}
# asymptotic covariance matrix
pander(cov2cor(est.covmat))
```

## residuals for 2-2 Tmers
```{r resids_2_2, echo=FALSE, results="asis", fig.align='center'}
resids <- counts_table(in_longwin=2, in_shortwin=2, half_n_show=20)
counts_plot(resids)
```

## residuals for 3-1 Tmers
```{r resids_3_1, results="asis", fig.align='center'}
resids <- counts_table(in_longwin=3, in_shortwin=1, half_n_show=20)
counts_plot(resids)
```

## residuals for 3-3 Tmers
```{r resids_3_3, results="asis", fig.align='center'}
resids <- counts_table(in_longwin=3, in_shortwin=3, half_n_show=20)
counts_plot(resids)
```

## residuals for 5-5 Tmers
```{r resids_5_5, results="asis", fig.align='center'}
resids <- counts_table(in_longwin=5, in_shortwin=5, half_n_show=20)
counts_plot(resids)
```

# Trace plot of the MCMC run:
```{r mcmc_trace, results="asis", echo=FALSE, fig.width=12, fig.height=7, fig.align='center'}
if (class(model)=="contextMCMC") {
    subseq <- seq(1,nrow(model@results$batch),length.out=min(1000,nrow(model@results$batch)))
    x <- sweep( model@results$batch[subseq,], 2, timevec[model@results$use.par], "/" )
    colnames(x) <- rownames(mr_compare)[model@results$use.par]
    param_sizes <- floor(log10(abs(mr_compare$simulated)+1e-4)/2)
    param_sizes <- param_sizes[model@results$use.par]
    for (k in unique(param_sizes)) {
        matplot( model@results$blen*subseq, x[,param_sizes==k], type='l', col=1:5, lty=1:6, 
            xlab='number of MCMC iterations', ylab='parameter values' )
        abline(h=mr_compare$simulated[model@results$use.par], col=1:5, lty=1:6)
        legend("topright",legend=colnames(x)[param_sizes==k],col=1:5,lty=1:6)
    }
} else {
    cat("Not an MCMC run.")
}
```

# Pairwise marginal posteriors from MCMC, colored by time in MCMC run
```{r mcmc_posterior, results="asis", echo=FALSE, fig.width=12, fig.height=12, fig.align='center'}
if (class(model)=="contextMCMC") {
    subseq <- seq(1,nrow(model@results$batch),length.out=min(1000,nrow(model@results$batch)))
    x <- rbind( x, mr_compare$simulated[model@results$use.par] ) 
    # color by time
    cols <- c( adjustcolor(rainbow(64),.4)[1+floor(65*(1:(nrow(x)-1))/nrow(x))], adjustcolor("black",.8) )
    pairs( x, col=cols, pch=20, cex=c(rep(.5,nrow(x)-1),2), gap=0 )
} else {
    cat("Not an MCMC run.")
}
```