### 2015-08-04 likelihood-surface.Rmd
.

---
title: "The likelihood surface"
date: "`r date()`"
---

```{r setup, include=FALSE}
required.input <- c("modelfile","longwin","shortwin","leftwin","tlen","ncounts")
if (!all(sapply(required.input,exists))) {
    stop("Not all required input (", paste(required.input[!sapply(required.input,exists)],collapse=", "), ") has been given.")
}
names(required.input) <- required.input
opt <- lapply( required.input, get )
opt$boundary <- "none"
if (!exists("lik.grid")) { lik.grid <- 10 }
source("context-inference-fns.R",chdir=TRUE)
library(MASS)
library(numDeriv)
library(pander)
if (!file.exists(modelfile)) { stop("Can't find ", modelfile) }
model.config <- fill.default.config( read.config(modelfile) )
model.config$fixfn <- parse.fixfn( model.config$fixfn, model.config$fixfn.params )
if (!exists("gmfile")) {
    gmfile <- file.path(dirname(modelfile),gsub("%",opt$longwin,model.config$genmatrix,fixed=TRUE))
}
if (file.exists(gmfile)) {
    load(gmfile)
} else { 
    cat("Can't find generator matrix ", gmfile, " -- making a new one.\n")
    genmatrix <- do.call( makegenmatrix,
                    c( list(
                        patlen=opt$longwin, 
                        mutpats=model.config$mutpats, 
                        selpats=model.config$selpats, 
                        selfactors=model.config$selfactors, 
                        boundary=opt$boundary, 
                        bases=model.config$bases, 
                        fixfn=model.config$fixfn ),
                        model.config$fixfn.params
                    )
                )
}
genmatrix@x <- do.call( update, c( list(G=genmatrix, mutrates=model.config$mutrates, selcoef=model.config$selcoef), model.config$fixfn.params ) )
projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=opt$leftwin, rightwin=opt$longwin-opt$shortwin-opt$leftwin, bases=genmatrix@bases )
```

Here is an estimate of the likelihood surface near the true values for the model in `r modelfile`,
at tuple size:

*  *longwin* : `r opt$longwin`
*  *shortwin* : `r opt$shortwin`
*  *leftwin* : `r opt$leftwin`
and other parameters
*  *total time* : `r opt$tlen`  (`tlen`)
*  *number of counts* : `r opt$ncounts`  (`ncounts`)
*  *number of grid points for likelihood* : `r lik.grid` (`lik.grid`)

```{r expected_counts, include=FALSE}
expected.counts <- predictcounts(opt$longwin, opt$shortwin, opt$leftwin, initcounts=rep(ncounts/nrow(genmatrix),nrow(genmatrix)), genmatrix=genmatrix, projmatrix=projmatrix, tlen=opt$tlen)

initpar <- with(model.config, unlist(c(mutrates*opt$tlen,selcoef,fixfn.params)) )
parscale <- with(model.config, unlist( c(mutrates.scale*opt$tlen, selcoef.scale, fixfn.params.scale) ) )
names(initpar) <- names(parscale) <- c( mutnames(genmatrix@mutpats), selnames(genmatrix@selpats), fixparams(genmatrix) )

# skip these parameters
use.par <- ( parscale!=0 )
params <- initpar

likfun <- function (sub.params){
    # params are: mutrates, selcoef, fixparams
    params[use.par] <- sub.params
    fparams <- params[seq( 1+nmuts(genmatrix)+nsel(genmatrix), length.out=length(fixparams(genmatrix)) )]
    names(fparams) <- fixparams(genmatrix)
    genmatrix@x <- do.call( update, c( list( G=genmatrix,mutrates=params[1:nmuts(genmatrix)],selcoef=params[seq(1+nmuts(genmatrix),length.out=nsel(genmatrix))]), as.list(fparams) ) )
    # this is collapsed transition matrix
    subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=1, time="fixed") # shape=params[length(params)], time="gamma" )
    # return POSITIVE log-likelihood
    ans <- sum( counts@counts * log(subtransmatrix) )
    if (!is.finite(ans)) { print(paste("Warning: non-finite likelihood with params:",paste(params,collapse=", "))) }
    return(ans)
}

model <- new( "context",
             counts=expected.counts,
             genmatrix=genmatrix,
             projmatrix=projmatrix,
             mutrates=model.config$mutrates*opt$tlen,
             selcoef=model.config$selcoef,
             params=unlist(model.config$fixfn.params),
             results=list(use.par=use.par),
             likfun=likfun,
             invocation="by hand"
         )
expected.env <- new.env()
assign("counts",model@counts,envir=expected.env)
environment(model@likfun) <- expected.env

```

Model configuration (`r modelfile`)
===================================

```{r show_config, echo=FALSE}
pander(model.config)
```

Hessian of the likelihood
=========================

Here are the square roots of the diagonal of the inverse of the Hessian:

```{r hessian, echo=FALSE}
time <- as.numeric(opt$tlen)
timevec <- c( rep(as.numeric(opt$tlen),nmuts(model)), rep(1,length(coef(model))-nmuts(model)) )
h <- function(x)model@likfun(x*timevec[model@results$use.par])
likfun.hess <- hessian(h,method.args=list(eps=min(abs(coef(model))[1:nmuts(model)]/10)),x=coef(model)[model@results$use.par])
colnames( likfun.hess ) <- rownames( likfun.hess ) <- names(coef(model))[model@results$use.par]
# asymptotic covariance matrix
est.covmat <- (-1)*ginv(likfun.hess)
dimnames(est.covmat) <- dimnames(likfun.hess)
pander( sqrt(diag(est.covmat)) )
```
and the Hessian itself:
```{r hessian_itself, echo=FALSE}
pander(likfun.hess)
```

Profiles of the likelihood
==========================

And, the pairwise profiles of the likelihood:
```{r pairwise_prof, echo=FALSE}
likfun.grids <- likelihood.surface( model, tlen=opt$tlen, ngrid=lik.grid )
```