### 2017-04-19 writeups/check-likelihood.Rmd
tidied up

---
title: "Computing sequence likelihoods"
date: "`r date()`"
---

\newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
context.dir <- "context"
source(file.path(context.dir,"context-inference-fns.R"),chdir=TRUE)
source(file.path(context.dir,"sim-context-fns.R"),chdir=TRUE)
```

The basic result is that if $L_k$ is the log-likelihood of all the $(k,w)$ T-mers in a sequence,
and $w$ and $k$ are long enough,
then $L_{k+1}-L_k$ should equal the likelihood of the entire sequence.

Let's check this.

Computing the likelihood
========================

Here's a function to compute the likelihood.
Note that the full likelihood of the *counts* includes a combinatorial factor
for where the T-mers occur,
but we *don't want* that factor.
```{r lhood_fun, cache=TRUE}
# opt will be a list with 'longwin', 'shortwin', and 'leftwin' defined.

likelihood_fun <- function (opt,config,seq) {
    ## count-seq: count T-mers
    longpats <- getpatterns(opt$longwin,config$bases)
    shortpats <- getpatterns(opt$shortwin,config$bases)
    counts <- counttrans.list( list(longpats,shortpats), 
        seqlist=seq[c("initseq","finalseq")],
        leftwin=opt$leftwin, bases=config$bases, 
        shift=0, cyclic=TRUE )

    ## make-genmatrix
    genmatrix <- do.call( makegenmatrix, c( list( patlen=opt$longwin ), boundary="none", config ) )
    projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=leftwin(counts), fpatterns=colnames(counts) )
    # this is collapsed transition matrix
    subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=seq$tlen, time="fixed")
    # return log-likelihood (*not* negative log-likelihood)
    ans <- sum( counts@counts * log(subtransmatrix) )
    dummy <- function() { 1 } # this will allow us access back to this environment

    return(list(loglik=ans,dummy=dummy))
}
```

The simplest model
==================

Let's start with a model without context-dependence.
```{r acgt_model, cache=TRUE}
acgt.config.json <- '{
    "bases" : [ "A", "C", "G", "T" ],
    "initfreqs" : [ 0.25, 0.25, 0.25, 0.25 ],
    "mutpats" : [
        [ [ "A", "T" ] ],
        [ [ "T", "A" ] ],
        [ [ "C", "G" ] ],
        [ [ "G", "C" ] ],
        [ [ "A", "C" ] ],
        [ [ "T", "G" ] ],
        [ [ "A", "G" ] ],
        [ [ "T", "C" ] ],
        [ [ "C", "A" ] ],
        [ [ "G", "T" ] ],
        [ [ "C", "T" ] ],
        [ [ "G", "A" ] ]
    ],
    "mutrates" : [ 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0 ]
}'
acgt.config <- fill.default.config(read.config(json=acgt.config.json))
```

Long time
---------

First, we'll do it for a long time, so the distribution should be uniform:
```{r first_likelihood, cache=TRUE, dependson=c("acgt_model","lhood_fun")}
seq <- simseq.config( seqlen=100, tlen=10.0, config=acgt.config )
opts.numerator <- list( 
            list(longwin=4, shortwin=2, leftwin=1),
            list(longwin=5, shortwin=3, leftwin=1),
            list(longwin=6, shortwin=4, leftwin=1)
        )
opts.denominator <- lapply( opts.numerator, function (x) { x$shortwin <- x$shortwin-1; x } )
loglik.numerators <- loglik.denominators <- list()
for (k in seq_along(opts.numerator)) {
    loglik.numerators[[k]] <- likelihood_fun( opts.numerator[[k]], config=acgt.config, seq=seq )
    loglik.denominators[[k]] <- likelihood_fun( opts.denominator[[k]], config=acgt.config, seq=seq )
}
results <- data.frame( 
      longwin=sapply(opts.numerator,"[[","longwin"),
      numerators=sapply(loglik.numerators,"[[","loglik"),
      denominators=sapply(loglik.denominators,"[[","loglik"),
      theory=(-1)*length(seq[["initseq"]])*log(4)
      )
results$est_loglik <- results$numerators-results$denominators
results
```
The likelihood differences should be close to the theory.
Note that we're treating the string as cyclic, so we don't need to worry about edge effects.

Less long time
--------------

For a shorter time, since the total time is `r seq$tlen` and the mutation rate is 1.0 to each of the possible three other bases,
a base is randomized after an Exponential(4) time;
so the likelihood, where $m$ bases differ from the original and $n-m$ remain the same,
$$\begin{aligned}
\left( e^{-4t} + (1-e^{-4t}) \frac{1}{4} \right)^{n-m}
\left( (1-e^{-4t}) \frac{1}{4} \right)^m  .
\end{aligned}$$
```{r shorter_likelihood, cache=TRUE, dependson=c("acgt_model","lhood_fun")}
seq <- simseq.config( seqlen=100, tlen=0.1, config=acgt.config )
theoretical <- function (seq) {
    n <- length(seq$initseq)
    m <- stringDist( DNAStringSet( seq[c("initseq","finalseq")]), method="hamming" )
    m * ( log((1-exp(-4*seq$tlen))/4) ) + (n-m) * log( exp(-4*seq$tlen) - expm1(-4*seq$tlen)/4)
}
opts.numerator <- list( 
            list(longwin=4, shortwin=2, leftwin=1),
            list(longwin=5, shortwin=3, leftwin=1),
            list(longwin=6, shortwin=4, leftwin=1)
        )
opts.denominator <- lapply( opts.numerator, function (x) { x$shortwin <- x$shortwin-1; x } )
loglik.numerators <- loglik.denominators <- list()
for (k in seq_along(opts.numerator)) {
    loglik.numerators[[k]] <- likelihood_fun( opts.numerator[[k]], config=acgt.config, seq=seq )
    loglik.denominators[[k]] <- likelihood_fun( opts.denominator[[k]], config=acgt.config, seq=seq )
}
results <- data.frame( 
      longwin=sapply(opts.numerator,"[[","longwin"),
      numerators=sapply(loglik.numerators,"[[","loglik"),
      denominators=sapply(loglik.denominators,"[[","loglik"),
      theory=as.numeric(theoretical(seq))
      )
results$est_loglik <- results$numerators-results$denominators
results
```
The "diff" column should match the "theory" column.

Simple, context-dependent model
===============================

Now, let's check this looks sensible for a simple context-dependent one.
```{r binary_model, cache=TRUE, dependson=c("lhood_fun")}
xo.config.json <- '{
    "bases" : [ "O", "X" ], 
    "initfreqs" : [ 0.5, 0.5 ],
    "mutpats" : [
        [ [ "O", "X" ], [ "X", "O" ] ],
        [ [ "OO", "XX" ], [ "XX", "OO" ] ]
    ],
    "mutrates" : [ 1.0, 1.0 ]
}'
xo.config <- fill.default.config(read.config(json=xo.config.json))
xo.seq <- simseq.config( seqlen=10000, tlen=0.1, config=xo.config )
opts.numerator <- list( 
            list(longwin=4, shortwin=2, leftwin=1),
            list(longwin=6, shortwin=2, leftwin=2),
            list(longwin=8, shortwin=2, leftwin=3),
            list(longwin=5, shortwin=3, leftwin=1),
            list(longwin=7, shortwin=3, leftwin=2),
            list(longwin=8, shortwin=3, leftwin=3),
            list(longwin=9, shortwin=3, leftwin=3),
            list(longwin=6, shortwin=4, leftwin=1),
            list(longwin=8, shortwin=4, leftwin=2)
        )
opts.denominator <- lapply( opts.numerator, function (x) { x$shortwin <- x$shortwin-1; x } )
loglik.numerators <- loglik.denominators <- list()
for (k in seq_along(opts.numerator)) {
    loglik.numerators[[k]] <- likelihood_fun( opts.numerator[[k]], config=xo.config, seq=xo.seq )
    loglik.denominators[[k]] <- likelihood_fun( opts.denominator[[k]], config=xo.config, seq=xo.seq )
}
results <- data.frame( 
      longwin=sapply(opts.numerator,"[[","longwin"),
      shortwin=sapply(opts.numerator,"[[","shortwin"),
      numerators=sapply(loglik.numerators,"[[","loglik"),
      denominators=sapply(loglik.denominators,"[[","loglik")
      )
results$est_loglik <- results$numerators-results$denominators
results
```

With a prior on the branch length
========================

Now, let's modify the function to allow for a Gamma prior on the branch length,
with shape $\alpha$ and scale $\theta$.
```{r lhood_fun_gamma, cache=TRUE}
# opt will be a list with 'longwin', 'shortwin', and 'leftwin' defined.
likelihood_fun_gamma <- function (opt,config,seq,shape,scale) {
    ## count-seq: count T-mers
    longpats <- getpatterns(opt$longwin,config$bases)
    shortpats <- getpatterns(opt$shortwin,config$bases)
    counts <- counttrans.list( list(longpats,shortpats), 
        seqlist=seq[c("initseq","finalseq")],
        leftwin=opt$leftwin, bases=config$bases, 
        shift=0, cyclic=TRUE )
    ## make-genmatrix
    genmatrix <- do.call( makegenmatrix, c( list( patlen=opt$longwin ), boundary="none", config ) )
    projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=leftwin(counts), fpatterns=colnames(counts) )
    # this is collapsed transition matrix
    subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=scale, shape=shape, time="gamma")
    # return log-likelihood (*not* negative log-likelihood)
    ans <- sum( counts@counts * log(subtransmatrix) )
    dummy <- function() { 1 } # this will allow us access back to this environment
    return(list(loglik=ans,dummy=dummy))
}
```

The first model
--------------

In the first model, in the expression for the likelihood
we just need to adjust the probability of having mutated or not;
the probability that a Gamma($k$,$\theta$) occurs before a Exponential(4)
is
$$\begin{aligned}
\int_0^\infty \frac{1}{\Gamma(k)\theta^k} t^{k-1} e^{-t/\theta} e^{-4t} dt
&=
\int_0^\infty \frac{1}{\Gamma(k)\theta^k} t^{k-1} e^{-t\frac{4\theta+1}{\theta}} dt \\
&=
\frac{1}{\Gamma(k)\theta^k} \left(\frac{\theta}{4\theta+1}\right)^k \int_0^\infty u^{k-1} e^{-u} du \\
&=
\frac{1}{(4\theta+1)^k}
\end{aligned}$$
(also the Laplace transform for a Gamma distribution)
and so the likelihood of a sequence with $m$ sites differing is, as above,
$$\begin{aligned}
\left( \frac{1}{(4θ+1)ᵏ} + \left(1-\frac{1}{(4\theta+1)ᵏ}\right) \frac{1}{4} \right)^{n-m}
\left( \left(1-\frac{1}{(4\theta+1)ᵏ}\right) \frac{1}{4} \right)^m  .
\end{aligned}$$
```{r gamma_likelihood, cache=TRUE, dependson=c("acgt_model","lhood_fun_gamma")}
shape <- 3
scale <- 0.1/3

seq <- simseq.config( seqlen=100, tlen=0.1, config=acgt.config )
theoretical_gamma <- function (seq,shape,scale) {
    n <- length(seq$initseq)
    m <- stringDist( DNAStringSet( seq[c("initseq","finalseq")]), method="hamming" )
    p <- 1/(4*scale+1)^shape
    m * ( log((1-p)/4) ) + (n-m) * log( p + (1-p)/4 )
}
opts.numerator <- list( 
            list(longwin=4, shortwin=2, leftwin=1),
            list(longwin=5, shortwin=3, leftwin=1),
            list(longwin=6, shortwin=4, leftwin=1)
        )
opts.denominator <- lapply( opts.numerator, function (x) { x$shortwin <- x$shortwin-1; x } )
loglik.numerators <- loglik.denominators <- list()
for (k in seq_along(opts.numerator)) {
    loglik.numerators[[k]] <- likelihood_fun_gamma( opts.numerator[[k]], config=acgt.config, seq=seq, shape=shape, scale=scale )
    loglik.denominators[[k]] <- likelihood_fun_gamma( opts.denominator[[k]], config=acgt.config, seq=seq, shape=shape, scale=scale )
}
results <- data.frame( 
      longwin=sapply(opts.numerator,"[[","longwin"),
      shortwin=sapply(opts.numerator,"[[","shortwin"),
      numerators=sapply(loglik.numerators,"[[","loglik"),
      denominators=sapply(loglik.denominators,"[[","loglik"),
      theory=as.numeric(theoretical_gamma(seq, shape=shape, scale=scale))
      )
results$est_loglik <- results$numerators-results$denominators
results
```
The "diff" column should match the "theory" column.

Finding the MLE for the time
============================

We'll do this with a lightly context-dependent model.
```{r cpg_model, cache=TRUE}
cpg.config.json <- '{
    "bases" : [ "A", "C", "G", "T" ],
    "initfreqs" : [ 0.25, 0.25, 0.25, 0.25 ],
    "mutpats" : [
        [ [ "A", "T" ] ],
        [ [ "T", "A" ] ],
        [ [ "C", "G" ] ],
        [ [ "G", "C" ] ],
        [ [ "A", "C" ] ],
        [ [ "T", "G" ] ],
        [ [ "A", "G" ] ],
        [ [ "T", "C" ] ],
        [ [ "C", "A" ] ],
        [ [ "G", "T" ] ],
        [ [ "C", "T" ] ],
        [ [ "G", "A" ] ],
        [ [ "CG", "CT" ], [ "CG", "AG" ] ]
    ],
    "mutrates" : [ 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 0.8 ]
}'
cpg.config <- fill.default.config(read.config(json=cpg.config.json))
```

Fixed times
-----------

Let $P_t = e^{tG}$.
This involves computing $P_t V$, where $V$ is a projection matrix.
If we've already computed this for $t$,
and want to check the likelihood for $t+s$, 
then we need $P_{t+s} V = P_s P_t V$ --
so, the previous sub-transition matrix $P_t V$ can be re-used for larger times.

```{r infer_fixed_time_fun, cache=TRUE}
t_lhoods <- function (opt, config, seq, dt, ndt) {
    ans <- numeric(ndt)
    longpats <- getpatterns(opt$longwin,config$bases)
    shortpats <- getpatterns(opt$shortwin,config$bases)
    counts <- counttrans.list( list(longpats,shortpats), 
        seqlist=seq[c("initseq","finalseq")],
        leftwin=opt$leftwin, bases=config$bases, 
        shift=0, cyclic=TRUE )
    genmatrix <- do.call( makegenmatrix, c( list( patlen=opt$longwin ), boundary="none", config ) )
    projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=leftwin(counts), fpatterns=colnames(counts) )
    subtransmatrix <- projmatrix
    for (k in 1:ndt) {
        subtransmatrix <- computetransmatrix( genmatrix, subtransmatrix, tlen=dt, time="fixed")  # add time dt
        # return log-likelihood (*not* negative log-likelihood)
        ans[k] <- sum( counts@counts * log(subtransmatrix) )
    }
    return(ans)
}
lower_opt <- function (x) { x$shortwin <- (x$shortwin-1); x }
```

OK, here's the likelihood profile:
```{r infer_fixed_time, cache=TRUE, dependson=c("cpg_model","infer_fixed_time_fun")}
cpg.seq <- simseq.config( seqlen=10000, tlen=0.1, config=cpg.config )
opts <- list( 
            list(longwin=4, shortwin=2, leftwin=1)
        )
dt <- .005
ndt <- ceiling(.2/dt)
tt <- (1:ndt)*dt
numerators <- t_lhoods(opts[[1]], cpg.config, cpg.seq, dt, ndt )
denominators <- t_lhoods(lower_opt(opts[[1]]), cpg.config, cpg.seq, dt, ndt )
cbind(tlen=seq$tlen,num=numerators,den=denominators,diff=numerators-denominators)
plot( tt, numerators-denominators )
abline(v=seq$tlen)
```

Exponential times
-----------------

Consider the case $k=1$, i.e., Exponential.
Let $\lambda = \max_i -G_{ii}$ and $Q_{ij} = G_{ij}/\lambda$ and $Q_{ii}=-\sum_{j \neq i} G_{ij}$,
so that the process jumps as $Q$ at the times of a Poisson($\lambda$) process.
As described in [context/writeups/gamma-prob.html](the context project),
to move from an Exponential($1/\theta$) time to an Exponential($1/(\theta+u)$) time,
we need to add either zero jumps (with probability $\theta/(\theta+u)$)
or a Geometric($\lambda(\theta+u)/(1+\lambda(\theta+u))$) number of jumps.
This is done by `extend.expAtv`;
a similar thing could be done for Gamma times with shape not equal to 1.
```{r infer_expl_time_fun, cache=TRUE}
e_lhoods <- function (opt, config, seq, dscale, ndscale) {
    ans <- numeric(ndscale)
    longpats <- getpatterns(opt$longwin,config$bases)
    shortpats <- getpatterns(opt$shortwin,config$bases)
    counts <- counttrans.list( list(longpats,shortpats), 
        seqlist=seq[c("initseq","finalseq")],
        leftwin=opt$leftwin, bases=config$bases, 
        shift=0, cyclic=TRUE )
    genmatrix <- do.call( makegenmatrix, c( list( patlen=opt$longwin ), boundary="none", config ) )
    projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=leftwin(counts), fpatterns=colnames(counts) )
    # subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=dscale, time="gamma")  # add scale dscale
    subtransmatrix <- projmatrix
    scale <- 0
    for (k in 1:ndscale) {
        subtransmatrix <- extend.expAtv( A=genmatrix, scale=scale, dscale=dscale, gAtv=subtransmatrix )  # add scale dscale
        # return log-likelihood (*not* negative log-likelihood)
        ans[k] <- sum( counts@counts * log(subtransmatrix) )
    }
    return(ans)
}
lower_opt <- function (x) { x$shortwin <- (x$shortwin-1); x }
```

OK, here's the likelihood profile:
```{r infer_expl_time, cache=TRUE, dependson=c("cpg_model","infer_expl_time_fun")}
scale <- 0.1
shape <- 1
nseqs <- 1000
seqlen <- 100
random.tlen <- rgamma(nseqs,scale=scale,shape=shape)
expl.seqs <- simseq.config( seqlen=seqlen, tlen=random.tlen, config=cpg.config )
expl.seq <- list(
        initseq = do.call( xscat, lapply(expl.seqs,"[[","initseq") ),
        finalseq = do.call( xscat, lapply(expl.seqs,"[[","finalseq") ),
        scale=scale,
        shape=shape
    )
opts <- list( 
            list(longwin=4, shortwin=2, leftwin=1)
        )
dscale <- .005
ndscale <- ceiling(.2/dscale)
scalevals <- (1:ndscale)*dscale
numerators <- e_lhoods(opts[[1]], cpg.config, expl.seq, dscale, ndscale )
denominators <- e_lhoods(lower_opt(opts[[1]]), cpg.config, expl.seq, dscale, ndscale )
cbind(scale=scalevals,num=numerators,den=denominators,diff=numerators-denominators)
plot( scalevals, numerators-denominators, main=paste( names(opts[[1]]), opts[[1]], sep="=", collapse=", ") )
abline(v=expl.seq$scale)
```

Hm, let's try that again with a longer window?
```{r infer_expl_longer, cache=TRUE, dependson="infer_expl_time"}
longer.opt <- list( longwin=7, shortwin=3, leftwin=2 )
numerators <- e_lhoods(longer.opt, cpg.config, expl.seq, dscale, ndscale )
denominators <- e_lhoods(lower_opt(longer.opt), cpg.config, expl.seq, dscale, ndscale )
cbind(scale=scalevals,num=numerators,den=denominators,diff=numerators-denominators)
plot( scalevals, numerators-denominators, main=paste( names(longer.opt), longer.opt, sep="=", collapse=", ") )
abline(v=expl.seq$scale)
```

Not useful or not used
======================

Gamma inference
---------------

Let $f(t;k,\theta)$ be the density function for a Gamma distribution
with shape $k$ and scale $\theta$ (and hence mean $k \theta$).
Define
$$  P_{k,\theta} = \int_0^\infty f(t;k,\theta) P_t dt ,$$
i.e., the transition matrix after a Gamma($k$,$\theta$)-distributed time.
If $T_i$ are independent, Gamma($k_i$,$\theta$), then $\sum_i T_i$ is Gamma($\sum_i k_i$,$\theta$),
and so
$$\begin{aligned}
    P_{k_1,\theta} P_{k_2,\theta} &= P_{k_1+k_2,\theta} .
\end{aligned}$$

For $k$ an integer,
since a Gamma($k$,$\theta$) is the sum of $k$ independent Exponential($1/\theta$) times,
and by Poisson thinning,
a Gamma($k$,$\theta/(1-\epsilon)$) is the sum of a NegativeBinomial($k$,$1-\epsilon$) number of Exponential($1/\theta$) times,
i.e., the number of times to flip a $1-\epsilon$ coin until getting $k$ heads,
so
$$\begin{aligned}
    P_{k,\theta/(1-\epsilon)}
    &=
    \sum_{n \ge 0} 
    \binom{n+k-1}{n} \epsilon^{n} (1-\epsilon)^k
    P_{1,\theta}^{n+k}  .
\end{aligned}$$
Or, to get a Gamma($k$,$\theta/(1-\epsilon)$),
we could start with a Gamma($k$,$\theta$),
and add on $M$ indepenent Exponential($1/\theta$) times,
where $M$ is NegativeBinomial($N$,$1-\epsilon$)
and $N$ is Binomial($k$,$\epsilon$).

Extending gamma times
---------------------

Now consider the case $k=1$.
By Poisson thinning,
a Geometric($p$) number of independent Exponential($\lambda$) times
is Exponential($\lambda p$);
and therefore
$$\begin{aligned}
    P_{1,\theta/(1-\epsilon)}
    &=
    \sum_{n \ge 1} 
    \epsilon^{n-1} (1-\epsilon)
    P_{1,\theta}^n \\
    &=
    \sum_{n=1}^k
    \epsilon^{n-1} (1-\epsilon)
    P_{1,\theta}^n
    + O(\epsilon^k) .
\end{aligned}$$

```{r build_exponential_fun, cache=TRUE}
extend_transmatrix <- function (P,a,da,eps=da/(a+da),tol=1e-8) {
    # if P is e^{TG} where T is Exp(a)
    #   return e^{SG} where S is Exp(a+da), with da>0.
    # as above, a+da = a/(1-eps), so eps=(da/a)/(1+da/a)
    ans <- Pm <- (1-eps)*P
    m <- ceiling(log(tol)/log(eps))
    # cat("m=",m,"\n")
    for (k in seq(2,length.out=(m-1))) {
        Pm <- eps * P %*% Pm
        ans <- ans + Pm
    }
    return(ans)
}
# prob an exponential(theta) is greater than an exponential(1) is 1/(1+theta), so:
stopifnot( abs( 1/(1+0.25/.8) - extend_transmatrix( 1/(1+0.25), eps=0.2 ) ) <= 1e-8 )
stopifnot( abs( 1/(1+0.25/.8) - extend_transmatrix( 1/(1+0.25), eps=0.2, tol=1e-15 ) ) <= 1e-15 )
```

Here's a quick check of this:
```{r check_exponential_additivity, cache=TRUE, dependson="build_exponential_fun"}
## one step
scale <- 0.3
genmatrix <- do.call( makegenmatrix, c( list( patlen=2 ), boundary="none", acgt.config ) )
subtransmatrix <- computetransmatrix( genmatrix, Diagonal(nrow(genmatrix)), tlen=scale, shape=1, time="gamma", tol=1e-12 )
stm1 <- extend_transmatrix( subtransmatrix, eps=.2, tol=1e-12 )
stm2 <- computetransmatrix( genmatrix, Diagonal(nrow(genmatrix)), tlen=scale/0.8, shape=1, time="gamma", tol=1e-12 )
range(stm1-stm2)
stopifnot(all(abs(stm1-stm2)<1e-12))
## more steps
tt <- seq(0,scale,length.out=11)[-1]
stm.list <- vector(length(tt),mode='list')
stm.list[[1]] <- computetransmatrix( genmatrix, Diagonal(nrow(genmatrix)), tlen=tt[1], shape=1, time="gamma", tol=1e-12 ) 
for (k in seq_along(tt)[-1]) {
    stm.list[[k]] <- extend_transmatrix( stm.list[[k-1]], a=tt[k-1], da=diff(tt)[k-1], tol=1e-12 )
}
range(subtransmatrix-stm.list[[length(tt)]])
stopifnot(all(abs(subtransmatrix-stm.list[[length(tt)]])<1e-12))
```

Now let's look at inference.
Hm, since we're actually dealing with $P_{k,\theta} v$ instead of $P_{k,\theta}$ itself,
maybe that's not too useful.