### 2021-12-09 writeups/notes.tex

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[hidelinks]{hyperref}
\usepackage{bm}
\usepackage{todonotes}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\calS}{\mathcal{S}}  % set of states
\newcommand{\calT}{\mathcal{T}}  % set of transition triples
\newcommand{\calP}{\mathcal{P}}  % set of selection pairs
\newcommand{\calM}{\mathcal{M}}  % set of potential mutations
\newcommand{\nA}{\mbox{A}}  % nucleotides:
\newcommand{\nC}{\mbox{C}}
\newcommand{\nG}{\mbox{G}}
\newcommand{\nT}{\mbox{T}}
\newcommand{\nS}{\mbox{S}} % C or G
\newcommand{\nW}{\mbox{W}} % A or T
\newcommand{\nR}{\mbox{R}} % purine, A or G
\newcommand{\nY}{\mbox{Y}} % pyrimidine, C or T
\newcommand{\join}{\oplus}  % concatenation
\newcommand{\st}{\colon}  % such that
\newcommand{\var}{\mathop{\mbox{var}}}
\newcommand{\cov}{\mathop{\mbox{cov}}}
\newcommand{\given}{\;\mid\;}
\newcommand{\like}{\mathcal L}
\newcommand{\loglike}{\ell}
\newcommand{\alike}{\widetilde{\like}}
\newcommand{\aloglike}{\widetilde{\loglike}}

\newcommand{\APOBEC}{\text{\tiny APOBEC}}
\newcommand{\AID}{\text{\tiny AID}}

\begin{document}

Here are some bits removed from the main text that might provide useful context.

%%%%%%%%
\subsection{Edge effects and cyclicization}

The choice to treat long sequences as cyclical is generally of little importance.
However, we have defined in \eqref{eqn:G_defn} the basic building block of our approximation, $G(m)$,
to be the transition matrix \emph{for a circular sequence} of length $m$, where $m$ may be short.
\peter{no we haven't}
This seems strange -- but, what else would we do at the edges?
We want to use $G(\ell+m+r)$ to compute the chance that a given $(\ell+m+r)$-sequence ends up with another given $m$-sequence in the middle;
and if $\ell$ and $r$ are big enough relative to $t$, this will be a good approximation regardless of what we do at the edges.
One answer would be to not allow patterns hanging off of either end to match,
resulting in a slight underestimate of the chance of change.
A better approximation would be to use the marginal frequencies in some way to average over surrounding sequence
-- but, this requires marginal frequencies to be known, and these are not stationary.
However, the leftmost site in a randomly chosen window is a sample from the marginal distribution,
and if sites separated by distance $\ell+m+r$ are sufficiently uncorrelated,
wrapping the sequence around is a good approximation to averaging over marginal frequencies in the surrounding sequence.
It is easy to imagine situations where this would not be the best solution --
but again, if $\ell$ and $r$ are large enough, it won't matter much.

%%%%
\subsection{The root distribution}

We clearly don't want to add an extra parameter $\pi(z)$ for all $z \in \calS^{m+2\ell}$.
If nothing else, it would make the inference depend on $m$ and $\ell$, which we'd like to avoid.
One approach would be to put a Dirichlet distribution on $\pi(z)$.
This has the nice property that you can choose priors that are consistent across different values of $m$ and $\ell$.
We'd have to still MCMC around in the space of $\pi(z)$, so they'd be nuisance parameters from that point of view.
Also, making the frequencies at the root random introduces correlations between counts, which we want to avoid.
An alternative is to specify $\pi$ through lower-dimensional statistics,
for instance to model the root state as a sample from a Markov random field
with parameters for the marginal frequencies of each state, and pairwise (or higher) correlations between them.

A Markov random field with interactions between sites no more than distance two away
is determined by the energy function $\phi(a,b,c)$ which gives the ``energy'' of the configuration $abc$ for each $a,b,c \in \calS$.
We use this to define the distribution on root patterns $\pi$ by
\begin{align}  \label{eqn:root_mrf}
    \pi(z) = \frac{1}{Z} \exp\left( - \sum_{i=1}^{|z|} \phi(z_{i-1},z_i,z_{i+1}) \right),
\end{align}
i.e., the exponential of the total energy of each three adjacent sites,
where $Z$ is the normalizing constant.
Note that it is often difficult to compute $Z$ in applications of Markov random fields,
but here we are working with relatively short sequences,
so we just compute \eqref{eqn:root_mrf} for all possible $z$, and divide by the sum.
(However, for this same reason this is not equivalent to saying that $X(0)$ is a sample from the Markov random field
and finding the distribution of patterns within this.)

%%%%%
\subsubsection{Mixing}

There can be significant confounding of the frequencies at the root with the other parameters,
which results in poor mixing.
(See CpG results, below.)
To see why this is, consider the following back-of-the-envelope calculation.
Take $x \neq y$ to be single bases (so, set $\ell=0$ and $w=1$),
so that $N(x,y)$ gives the number of times an $x$ and a $y$ are homologous.
Let $q(x,y)$ be the instantaneous rate that $x \to y$ (averaged somehow over context?),
so that $N(x,y)$ should be, roughly,
\begin{align*}
  \sum_z \pi(z) q(z,x) q(z,y) &\approx t \pi(x) (1+q(x,x)) q(x,y) + t \pi(y) q(y,x) (1+q(y,y)) ,
\end{align*}
where as usual $q(x,x) = (-1) \sum_{z \neq x} q(x,z)$ is (minus) the total jump rate out of $x$.

So, when we propose a change to $\pi$ we might want to also adjust the rates accordingly to keep this more-or-less constant:
if we chnage $\pi(x) \mapsto \pi(x) + \epsilon_x$, we want $\delta_{xy}$ to satisfy
\begin{align*}
  \sum_z \pi(z) q(z,x) q(z,y) = \sum_z (\pi(z) + \epsilon_z) ( q(z,x) + \delta_{zx} ) ( q(z,y) + \delta_{zy} ) ,
\end{align*}
or, to first order in $\delta$,
\begin{align*}
  0 &\approx \sum_z \epsilon_z q(z,x) q(z,y) + \sum_z \pi(z) \delta_{zx} q(z,y) + \sum_z \pi(z) q(z,x) \delta_{zy} \\
   &\approx \sum_z \epsilon_z q(z,x) q(z,y) + \sum_z (\pi(z)+\epsilon_z) \delta_{zx} q(z,y) + \sum_z (\pi(z)+\epsilon_z) q(z,x) \delta_{zy} \\
   &\approx \epsilon_x (1+q(x,x)) q(x,y) + \epsilon_y q(y,x) (1+q(y,y)) + (\pi(y)+\epsilon_y) \delta_{yx} (1+q(y,y)) + (\pi(x)+\epsilon_x) (1+q(x,x)) \delta_{xy} \\
   &= \left( \epsilon_x q(x,y) + (\pi(x)+\epsilon_x) \delta_{xy} \right)(1+q(x,x)) + \left( \epsilon_y q(y,x) + (\pi(y)+\epsilon_y) \delta_{yx} \right)(1+q(y,y))
\end{align*}
One solution to this is to let $\delta_{xy} = - \epsilon_x q(x,y) / \pi(x)$.

Suppose each change is a single-base change, so that changes $\delta \mu_i$ in $\mu$ contribute to $q$ by
$\delta_{xy} = \sum_{i: x \to y} \delta \mu_i$.
Given $\delta_{xy}$, this gives a linear equation for $\delta \mu_i$ that may or may not be solvable,
but certainly is if we have all the single-base transitions as parameters.

\end{document}